% mnras_template.tex 
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
%\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}


%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{multirow}
\usepackage{txfonts}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{microtype}

\usepackage{subfiles} % to split the paper into different files
\usepackage{xcolor}
\usepackage[decimalsymbol=., expproduct=times, separate-uncertainty=true, multi-part-units=single]{siunitx} % SI units 
\DeclareSIUnit\erg{erg}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand\note[1]{{\colorbox{yellow!60}{\color{magenta}#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[External photoevaporation of disks in young star clusters]{External photoevaporation shapes the distributions of circumstellar disk sizes in young star clusters}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Concha-Ramírez et al.]{
Francisca Concha-Ramírez\thanks{E-mail: fconcha@strw.leidenuniv.nl},
Martijn Wilhelm,
Simon Portegies Zwart
\\
% List of institutions
% List of institutions
%$^{1}$
Leiden Observatory, Leiden University, PO Box 9513, 2300 RA Leiden, The Netherlands\\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2018}

% Don't change these lines
\hypersetup{draft}
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
Circumstellar disks develop shortly after star formation, and during their first million years of evolution they are immersed in surroundings with high gas and stellar densities. This environment is unfavorable for disk survival: they can be truncated by dynamical encounters, evaporated by the radiation of bright OB stars nearby, shrunk by ram pressure stripping, and diminished by stellar winds and supernovae explosions. Circumstellar disks are also subject to viscous spreading, a consequence of angular momentum conservation. In the present work we used numerical and semi-analytical approaches to quantify the effects of dynamical truncations, external photoevaporation, and viscous evolution of circumstellar disks inside star clusters. We ran simulations of stellar clusters with $N=$\note{X} stars in a Plummer sphere with virial radius \note{Y pc}. We evolved these clusters for \note{Z} Myr, and analyzed the mean disk lifetimes at the end of the simulations. \note{We find that...} testing overleaf github stuff
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
key 1 key 2
\end{keywords}

\section{Introduction}
\label{sec:introduction}

Circumstellar disks develop shortly after star formation \citep{williams2011}. Young star forming regions, less than $\SI{2}{Myr}$ old, have been observed to have protoplanetary disks in more than 80\% of the cases \citep{fedele2010}. Since most stars are not born in isolation, and gas left over from the star formation process can linger for a few Myr \citep{goodwin2009,portegieszwart2010}, the disks remain embedded in this environment during their first stages of evolution. These conditions can be hostile for the disks in a myriad of ways. The surrounding gas can shrink the disks through ram pressure stripping and face-on accretion \citep{wijnen2016, wijnen2017}. Regarding the stellar dense environment, disks can be subject to dynamical truncations \citep{portegieszwart2016, vincke2015, vincke2016} or be affected by processes related to stellar evolution, such as stellar winds \citep{pelupessy2012}, supernovae explosions \citep{close2017}, and photoevaporation due to bright OB stars in the vicinity \citep[e.g.][]{haworth2017, guarcello2016}. Since planet formation related processes seem to start very quickly in circumstellar disks ($< 0.1 - \SI{1}{Myr}$, \citet{manara2018,najita2014}), understanding the mechanisms that affect disk evolution is directly connected with understanding planetary system formation and survival. Our own Sun was born within a star cluster \citep{adams2010,portegieszwart2009}, so discerning how the cluster environment affects the evolution of the disks can help us comprehend the origins of our home, the Solar System. 

There are observational indications that the environment in which the disks are surrounded just after their formation is unfavorable for their survival. Circumstellar disks have been observed to be evaporating in several star forming regions \cite[e.g.][]{mann2014,fang2012,dejuanovelar2012}. Moreover, observations suggest that disk fractions in cluster environments decline by a factor of $\sim2$ in regions close to an O-type star \citep{balog2007,guarcello2007,guarcello2009,fang2012,guarcello2016}. \citet{fatuzzo2008} estimated an FUV radiation field of up to $G_0 \approx 1000$ in star clusters of $N > 1000$ stars\footnote{$G_0$ is the FUV field in Habing units,  $\SI{1}{G_0} = \SI{1.6e-3}{\erg\per\second\per\square\cm}$. The solar vicinity is estimated to have a background FUV field of $G_0 = 1$ \citep{habing1968,parravano2003}.}, while \citet{facchini2016} showed that disks of sizes $\sim\SI{150}{au}$ are subject to photoevaporation even in very low FUV fields ($G_0 = 30$). Observational evidence of the imprints of dynamical truncations has been reported in regions of high stellar density \citep{olczak2008,reche2009,dejuanovelar2012}. Tidal stripping which can be explained by disk-star interaction has been observed in the RW Aurigae system \citep{cabrit2006,dai2015}. There is evidence that our own Solar System might have been affected by a close encounter with another star during its early stages \citep{pfalzner2018, jilkova2015}. It has also been observed that disks in dense regions are generally smaller than disks in more sparse regions \citep{mann2014,dejuanovelar2012} \note{i'm sure i have more references about this}, and it is still up for debate if this is a consequence of dynamical encounters or external photoevaporation, or a combination of both. 

Circumstellar disks are not only affected by external mechanisms, but also by their own viscous growth. The combination of outwardly decreasing angular velocity, together with outwardly increasing angular momentum, causes shearing forces inside the disks. These forces induce the transport of angular momentum from the inner regions, locally lower in angular momentum, to the outer parts, which have higher local angular momentum. As a consequence of this, mass is accreted from the innermost regions of the disk into its host star, while the outermost regions expand \citep{lynden-bell1974}. Observational reports of viscous spreading of disks around T Tauri stars have been described by \citet{isella2009} and \citet{guilloteau2011}. 

From a computational perspective, different approaches have been implemented to assemble these processes together and study their effects on the lifetimes of circumstellar disks. \note{photoevaporation...}. Analyses of the impact of dynamical truncations have been carried out in different ways. The effects of close encounters on a single SPH disk have been analyzed in several studies \note{citation, pfalzner, winter}. Another approach, in which the dynamics of the cluster are evolved separately, and then the effects of encounters are quantizied over an SPH disks, have also been published \note{pfalzner + vincke}. Directly adding SPH disks to a simulation of a massive star cluster is extremly computationally expensive. The closest effort corresponds to the recent work by \citet{rosotti2014}. Individual SPH codes were coupled to each one of the $\SI{1}{M_\odot}$ stars in a small ($N=100$) cluster. This allowed them to study the effects of viscous spreading of the disks and dynamical truncations in a self-consistent way, but were clearly limited by the computational resources needed for this problem. Parametrized \citep[e.g.][]{concha-ramirez2019,portegieszwart2016}\note{add more} and statistical approaches \citep[e.g.][]{olczak2006,malmberg2011}\note{add more} have also been developed. 
Similar work has been carried out by \citet{adams2006,winter2018a} but they average the $G_0$ on cluster environments of different sizes and use this over an SPH disk were truncations are dealt with separately (using Monte Carlo for \citet{winter2018a}).

In \citet{concha-ramirez2019} we investigated the effect of viscous growth and dynamical truncations on protoplanetary disks inside stars clusters. Both mechanisms were implemented with parametrized models. We showed that considering only viscous evolution and dynamical encounters yields circumstellar disks too large to be comparable to the compact disks observed in star forming regions. We argued that other processes must be at play to produce the observed disk size, mass, and acretion rate distributions. In the present work we expand our model by improving our description of the viscous growth of the disks, and by adding external photevaporation to the processes that can shrink the disks. In this way we look to better represent the environment of star forming regions. 

Instead of the parametrized description used in our previous work, we now modeled the circumstellar disks using the numerical code VADER \citep{krumholz2015}. VADER solves the equations of angular momentum and mass transport in a thin, axisymmetric disk, providing us with a numerical tool to implement the viscous expansion that is less computationally expensive than SPH disks. While the dynamical truncations were still parametrized, the mass loss due to external photevaporation was calculated using the FRIED grid \citep{haworth2018}. This grid consists of photoevaporative mass loss rates for disks of different sizes and masses, immersed in several different FUV fields. We used the Astrophysical Multipurpose Software Environment \citep[AMUSE\footnote{\url{http://amusecode.org}},][]{portegieszwart2013,pelupessy2013} framework to bring these codes together along with cluster dynamics and stellar evolution. All the code developed for the simulations, data analyses, and figures of this paper is available in a Github repository\footnote{ZENODO LINK}.


\section{Model}
\label{sec:model}

The circumstellar disks in our simulations were subject to viscous growth, dynamical truncations, and external photoevaporation. 

\subsection{Viscous growth of circumstellar disks}\label{viscous}
We implemented the protoplanetary disks with viscous evolution using the numerical code VADER (Viscous Accretion Disk Evolution Resource) by \citet{krumholz2015}. VADER is an open source viscous evolution code which solves the equations of mass transport, angular momentum transport, and internal energy in a thin, axisymmetric disk. An AMUSE interface for VADER was developed in the context of this work and is available online\note{add footnote w/ link -- check if Martijn is ok with this!}.  

Each VADER disk is composed of a grid of radial cells .... 

\note{add description of vader disks}

For our initial disk column density we used the standard disk profile introduced by \citet{lynden-bell1974}:

\begin{equation}\label{profile}
\Sigma(r, t=0) = \Sigma_0 \frac{r_c}{r} \exp\left(\frac{-r}{r_c}\right)
\end{equation}

\noindent
with

\begin{equation}
\Sigma_0 = \frac{m_d}{2 \pi {r_c}^2 \left(1 - \exp\left(-r_d/r_c\right)\right)}
\end{equation}

\noindent
where $r_c$ is the characteristic radius of the disk, $r_d$ and $m_d$ are the initial radius and mass of the disk, respectively, and $\Sigma_0$ is a normalization constant. Considering $r_c \approx r_d$ \citep{anderson2013}, the density profile of the disks takes the form:

\begin{equation}
\Sigma(r, t=0) \approx \frac{m_d}{2 \pi r_d \left(1 - \exp^{-1}\right)} \frac{\exp(-r/r_d)}{r}.
\end{equation}

This expression allows for the disk to expand freely at the outer boundary, while keeping the condition of zero torque at the inner boundary $r_i$ ($\Sigma(r_i)$ = 0).

To define the radius of our disks, we set the column density outside $r_d$ to a negligible value $\Sigma_{edge} = \SI{e-12}{\gram\per\square\cm}$. 

The temperature profile of our disks is given by

\begin{equation}
T(r) = T_m \left(\frac{r}{\SI{1}{au}}\right)^{-1/2}
\end{equation}

\noindent
where $T_m$ is the midplane temperature at $\SI{1}{au}$. Based on \citet{anderson2013} we adopt the value $T_m = \SI{300}{\K}$.


Each of our disks consisted of a grid of 500 logarithmically spaced cells, in a range between $0.05$ and $\SI{5000}{au}$, with a Keplerian rotation profile. The accreted mass that flows through the inner boundary was added to the mass of the host star.

\subsection{Dynamical truncations}\label{truncations}
Between encounters, the circumstellar disks evolved according to the numerical solutions of the transport equations used by VADER. A close encounter with another disk induces a discontinuity in this evolution. To modify the VADER disks we first calculated parametrized values for truncation radius and mass loss. \citet{rosotti2014} approximated the truncation radius to a third of the encounter distance for stars of the same mass. To this we added the mass dependence of \citet{bhandare2016}. Our truncation radius then takes the form:

\begin{equation}\label{truncationradius}
{r^\prime} = \frac{r_{enc}}{3}\left(\frac{M}{m}\right)^{0.5}
\end{equation}

\noindent 
where $M$ and $m$ are the masses of the encountering stars.

To implement disk truncations on the VADER disk codes, we first calculated the corresponding truncation radius for the encounter, according to equation \ref{truncationradius}. We then move outside-in through the disk cells and change the value of the column density to a lower value $\Sigma_{edge}$, as described in section \ref{truncations}. We stopped at the value corresponding to the truncation radius. In this way we defined a new disk edge at that position.

Close dynamical encounters not only affect the radius of the affected disks, but also their mass. We followed \citet{portegieszwart2016} to define the amount of mass loss from each disk as:

\begin{equation}
dm = m_{d}\frac{r^{1/2} - R^{1/2}}{r^{1/2}}
\end{equation}

\noindent
where $r$ and $R$ are the radii of the disks involved. We calculated the amount of mass that each disk accretes by:

\begin{equation}\label{accmass}
dm_{acc} = f \frac{m}{M + m} dm
\end{equation}

\noindent
where $f \leq 1$ is a mass transfer efficiency factor. For our calculations we used $f = 1$. Both equations of mass accretion were applied symmetrically to each of the disks involved in the close encounter.

\subsection{External photoevaporation}\label{photoevaporation}
To calculate the photoevaporative mass losses of the disks, we needed to obtain the mass loss ratio induced by the massive stars at different points of the simulation. For this we needed to first calculate the FUV luminosities of the massive stars at every time step, and then use this and other stellar values to obtain the amount of radiation that is reached by each disk. Below we explain how we performed each of these steps. 

\subsubsection{Photoevaporative mass loss}\label{grid}
To calculate the mass loss due to photoevaporation on each disk, we used the FRIED (Far-ultraviolet Radiation Induced Evaporation of Discs) grid developed by \citet{haworth2018}. The FRIED grid is an open access grid of calculations of externally evaporating circumstellar disks. It spans a six dimension parameter space consisting of disk sizes ($1 - 400 au$), disk masses ($~10^{-4} - 10^{2} M_{Jup}$), UV fields ($10 - 10^{4} G_0$), stellar masses ($0.05 - 1.9 M_{\odot}$) and disk outer surface densities. The seemingly
three dimensional subspace of disk mass, edge density and disk radius is in fact two dimensional, as any combination of disk radius and disk mass has only one edge density associated with it. Because of this, we only took into account a four dimensional grid of radiation field strength, host star mass, disk mass and disk radius.

Following the stellar mass ranges of the FRIED grid, we separated the stars in our clusters in two subgroups: \textit{bright stars} and \textit{small stars}. Bright stars are all stars with initial masses higher than $1.9 M_{\odot}$, while small stars have masses lower or equal than $1.9 M_{\odot}$. Only the small stars will have disks around them. The big stars will be considered as only generating FUV radiation and affecting the disks around the smaller stars. In this way we made sure that we stayed within the stellar mass limits of the grid. Low mass stars ($\lesssim 1 M_\odot$) have a negligible UV flux \citep{adams2006}, so this approximation holds well for our purposes. Calculation of the FUV radiation emitted by the bright stars is further explained in section \ref{FUVluminosities}. In terms of gravitational evolution, both star groups were indistinguishable and they evolved together normally in the gravity code. These star subgroups were considered only for the calculation of FUV radiation and photoevaporative mass loss.

The FRIED grid allows to take a circumstellar disk with a specific mass, size, and density, around a star with a certain mass, and from the FUV radiation that it receives, obtain the photoevaporative mass loss. However, the parameters of our disks not always exactly matched the ones in the grid. We performed an interpolation over the grid to calculate the photoevaporative mass losses needed for our specific disks. Because of computational constraints, we performed the interpolations on a subspace of the grid, such that it contains at least one data point above and below the phase space point of the disk in each dimension.

\subsubsection{FUV luminosities}\label{FUVluminosities}
External photoevaporation of protoplanetary disks in proximity to OB-type stars is dominated by FUV radiation \citep{guarcello2016,storzer1999,gorti2016a}. We followed \citet{adams2004} in defining the FUV band ranging from 6 eV up to 13.6 eV, the Lyman break, or approximately from 912 Å to 2070 Å. Given the presence of spectral lines in this band, we chose to use synthetic stellar spectra rather than relying on black body approximations. Our spectral library of choice was UVBLUE \citep{rodriguez-merino2005}, for its high coverage of parameter space and high resolution, spanning our needed wavelength ranges. To make our computations less expensive we used versions of the spectra with a resolving power five times lower than the full ones, which was sufficient for our calculations. 

The UVBLUE spectral library spans a three dimensional parameter space of stellar temperature, metallicity, and surface gravity. We used solar metallicity for all our stars, with a value of \note{X}. For the temperature and surface gravity, we chose the spectra that most close resembled the zero age main sequence (ZAMS) values of each star, according to the parametrized stellar evolution code SeBa \citep{portegieszwart1996, toonen2012}. \note{should I add a more detailed explanation of this process?}

Using the ZAMS spectra, we calculated the flux and corresponding luminosity using the radius of the star. Given that we knew the masses of the stars, we built a relationship between stellar mass and FUV luminosity. This relation takes the shape of a segmented power law, as can be seen in Figure X. A similar fit was obtained by \citet{parravano2003}.
%Ask for more details of this process?

% FUV luminosity vs mass figure!

% Add table?

The mass range of our fit is $0.12 - 100 M_{\odot}$, however, we were only interested in the range $1.9 - 100 M_{\odot}$. As explained in section \ref{grid}, we only considered stars with masses higher that $1.9 M_{\odot}$ to be emitting FUV radiation, and $100 M_{\odot}$ was the upper limit for our stellar mass distribution. These stars were also subject to stellar evolution, which was implemented with SeBa through the AMUSE interface. The FUV luminosity of each bright star is calculated in every time step, after going through stellar evolution.

\subsubsection{Calculation of photoevaporative mass loss}\label{massloss}
As explained in section \ref{grid}, we separated the stars in our cluster into two subgroups: bright stars, which emit FUV radiation and do not have disks, and small stars, which have disks around them. As such, small stars are the only ones affected by external photoevaporation, induced by the FUV radiation of the bright stars.

To calculate the amount of mass loss for the disks around the small stars, the following process was carried out in every time step of our simulation:

\begin{enumerate}
\item For each bright star, we calculated its FUV luminosity $L_{FUV}$ 
\item For each small star, we calculated the distance to each bright star $d$
\item Using $L_{FUV}$ and $d$, we calculated the amount of FUV radiation, $\ell_{FUV}$, received by the small star \note{check for proper symbols}
\item Using $\ell_{FUV}$ and the stellar mass, disk mass, and disk size of the small star, we built a subgrid of the FRIED grid and interpolated on it to calculate $\dot{M}_FUV$, the photoevaporative mass loss 
\item We calculate the total photoevaporative mass loss for the time step for each small star, and subtract that amount of mass from the disk mass
\item The disk radius of each small star is shrunk depending on the amount of mass lost due to photoevaporation
\end{enumerate}

\subsection{Evolution of the disks}
For each simulation we did the following:

-Start a Plummer sphere with N number of stars and virial size N

-Select all the stars under 1.9 MSun (because of the FRIED grid parameter space). To all these stars we gave circumstellar disks.

-For all the stars > 1.9 MSun (a.k.a. massive stars) we implemented stellar evolution.

-We used the UVBLUE spectral library to calculate the FUV luminosity of each of these massive stars.

-On each timestep we calculate the positions of the stars, and the radiation that each small stars gets from all the massive stars, in the distance

-This gives us the total photoevaporative mass loss rate for each small star

-We update the size of the circumstellar disk according to the viscous growth and to the photoevaporative mass loss rate


We consider the disks to be completely evaporated when they have lost 99\% of their initial mass, based on \citet{anderson2013}.

\section{Results}
\label{sec:results}

\subsection{Initial conditions}
\subsubsection{Cluster}
We run X simulations with N stars each located in Plummer (King?) spheres of Z size. We run M simulations lasting for 10 million years each. 

We then plot the cumulative distributions of the protoplanetary disks. To calculate the size of the protoplanetary disks we take the mass inside a characteristic radius. We set this radius to 95\% to make it similar to observations (check reference on my previous paper).

\subsubsection{Disks}
The initial sizes of our circumstellar disks were given by:

\begin{equation}
r_d(t=0) = R'\left(\frac{M_*}{M_\odot}\right)^{0.5}
\end{equation}

\noindent
where $R'$ is a constant. The youngest circumstellar disks observed to date have sizes that range from $\sim\SI{30}{au}$ \citep[e.g.][]{lee2018} to $\sim120-\SI{180}{au}$ \citep[e.g.][]{murillo2013,vanthoff2018}. Based on this we chose the value $R' = \SI{100}{au}$, which for our mass range $0.05-1.9 M_\odot$ for stars with disks yields initial sizes between $\sim\SI{22}{au}$ and $\sim\SI{137}{au}$.

\subsection{Photoevaporation vs dynamical truncations}

\subsection{Disk lifetimes}

\subsection{Comparison with observations}
\note{or should this go in the discussion?}

\section{Discussion}
\label{sec:discussion}

\subsection{Caveats}
Our model encompassed a simplified implementation of the processes inside young star forming regions. Several mechanisms and circumstances were not considered for our calculations. Regarding photoevaporation, our model ignores the different regimes in which this can occur. External photoevaporation can ensue in super and subcritical regimes. These division is established in terms of the critical radius $r_g$ at which the sound speed of the gas equals the escape speed from the disk \citep{adams2004}:

\begin{equation}
r_g = \frac{GM_*\left\langle\mu\right\rangle}{kT} \approx \SI{100}{au} \left(\frac{T}{\SI{1000}{K}}\right)^{-1}\left(\frac{M_*}{\SI{1}{M_\odot}}\right).
\end{equation}

The photoevaporative flows are different for disks with radius $r_{disk} < r_g$ (subcritical) and $r_{disk} > r_g$ (supercritical). \citet{adams2004} showed that the mass outflow rates are different for each of these cases, and that many disks fall in the subcritical regime. They also showed that even for disks where photoevaporation is suppressed because they have reached the subcritical size, significant mass loss still takes place as long as $r_{disk}/r_g \gtrsim 0.1-0.2$. This difference in photoevaporation regimes was not considered in our implementation, where all mass is equally removed independent of the radius of the disk where is it located. Also, our implementation only considered photoevaporative mass loss as working from the outside in, and did not include mass that can be removed from the surface of the disk at smaller radii. 

Furthermore, our implementation completely ignored the photoevaporation caused by the radiation of the disk's host star. \note{elaborate} \citep{fatuzzo2008}

We also did not differentiate between FUV and EUV radiation. This approximation is valid since it is unlikely that EUV radiation drives the photoevaporative mass loss of disks \citep{owen2012}. FUV photons have a much larger penetrating depth than EUV, and the former generate an ionization front several radii away from the disk surface which absorbs the incident EUV flux \citep{adams2004}. 

-We ignore dust growth and ignore different components for dust and gas disks 

-We ignore binaries (cite cool paper on circumbinary stuff)

-We ignore gas in the star forming regions

Stellar evolution was used only as a way to obtain the FUV luminosity of the massive stars at each moment in the evolution of the clusters. Outcomes of stellar evolution, such as stellar winds and supernovae explosions, were not considered in the simulations, but they could have a significant effect on the development of circumstellar disks inside star clusters \citep{close2017,portegieszwart2018,pelupessy2012}.

Disk winds could also be an important cause of mass loss on the disks. While some hints of disk winds have been observed, mass loss rates related to this process are still unconstrained due to instrumental restrictions \citep{pascucci2018}. It is expected that new generations of instruments such as ngVLA \note{citation} will help determine how critical these winds can be to disk evolution.

\subsection{Comparison with observations?}

\section{Summary}
\label{sec:summary}


% Don't change these lines
\bsp	% typesetting comment
\bibliographystyle{mnras}
\bibliography{references}
\label{lastpage}
\end{document}

% End of mnras_template.tex
