% mnras_template.tex 
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
%\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}


%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{multirow}
\usepackage{txfonts}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{microtype}

\usepackage{subfiles} % to split the paper into different files
\usepackage{xcolor}
\usepackage[decimalsymbol=., expproduct=times, separate-uncertainty=true, multi-part-units=single]{siunitx} % SI units 
\DeclareSIUnit\erg{erg}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand\note[1]{{\colorbox{yellow!60}{\color{magenta}#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[External photoevaporation of disks in young star clusters]{External photoevaporation shapes the distributions of protoplanetary disks in young star clusters}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Concha-Ramírez et al.]{
Francisca Concha-Ramírez\thanks{E-mail: fconcha@strw.leidenuniv.nl},
Martijn Wilhelm,
Simon Portegies Zwart
\\
% List of institutions
% List of institutions
%$^{1}$
Leiden Observatory, Leiden University, PO Box 9513, 2300 RA Leiden, The Netherlands\\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

% Don't change these lines
%\hypersetup{draft}
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
Circumstellar disks develop as a result of the star formation process, and during their first million years of evolution they are immersed in surroundings with high gas and stellar densities. This environment is unfavorable for disk survival: they can be truncated by dynamical encounters, evaporated by the radiation of bright OB stars nearby, shrunk by ram pressure stripping, and diminished by stellar winds and supernovae explosions. Circumstellar disks are also subject to viscous spreading, a consequence of angular momentum conservation. In the present work we used numerical and semi-analytical approaches to quantify the effects of dynamical truncations, external photoevaporation, and viscous evolution of circumstellar disks inside star clusters. We ran simulations of stellar clusters with $N=$\note{X} stars in a Plummer sphere with virial radius \note{Y pc}. We evolved these clusters for \note{Z} Myr, and analyzed the mean disk lifetimes at the end of the simulations. \note{We find that...}
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
key 1 key 2
\end{keywords}

\section{Introduction}
\label{sec:introduction}

Circumstellar disks develop as a result of the star formation process \citep{williams2011}. Star forming regions less than $\SI{2}{Myr}$ old have been observed to have protoplanetary disks in $\sim60\%$ of their stars \citep{fedele2010}. Since at least a non negligible fraction of stars are not born in isolation \citep{bressert2010,king2012}, and gas left over from the star formation process can linger for a few Myr \citep{goodwin2009,portegieszwart2010}, the disks remain embedded in this environment during their first stages of evolution. These conditions can be hostile for the disks in a myriad of ways. The surrounding gas can shrink the disks through ram pressure stripping and face-on accretion \citep{wijnen2016, wijnen2017}. Regarding the dense stellar environment, disks can be subject to dynamical truncations \citep{vincke2015,portegieszwart2016,vincke2016} or be affected by processes related to stellar evolution, such as stellar winds \citep{pelupessy2012}, supernovae explosions \citep{close2017}, and photoevaporation due to bright OB stars in the vicinity \citep[e.g.][]{guarcello2016,haworth2017}. Since planet formation related processes seem to start very quickly in circumstellar disks ($< 0.1 - \SI{1}{Myr}$, \citet{najita2014,manara2018}), understanding the mechanisms that affect disk evolution is directly connected with understanding planetary system formation and survival. Our own Sun was born within a star cluster \citep{portegieszwart2009}, so discerning how the cluster environment affects the evolution of the disks can help us comprehend the origins of our home, the Solar System. 

There are observational indications that the environment in which the disks are surrounded just after their formation is unfavorable for their survival. Circumstellar disks have been observed to be evaporating in several star forming regions \cite[e.g.][]{fang2012,dejuanovelar2012,mann2014}. Moreover, observations suggest that disk fractions in cluster environments decline in regions close to an O-type star \citep[e.g.][]{balog2007,guarcello2007,guarcello2009,guarcello2010,fang2012,guarcello2016}. \citet{fatuzzo2008} estimated an FUV radiation field of up to $G_0 \approx 1000$ in star clusters of $N > 1000$ stars\footnote{$G_0$ is the FUV field measured in units of the Habing flux, $\SI{1.6e-3}{\erg\per\second\per\square\cm}$ \citep{habing1968}. The solar vicinity is estimated to have a background FUV field of around $G_0 = 1.6$ \citep{parravano2003}.}, while \citet{facchini2016} showed that disks of sizes $\sim\SI{150}{au}$ are subject to photoevaporation even in very low FUV fields ($G_0 = 30$). Observational evidence of the imprints of dynamical truncations has been reported in regions of high stellar density \citep{olczak2008,reche2009,dejuanovelar2012}. Tidal stripping that can be explained by disk-star interaction has been observed in the RW Aurigae system \citep{cabrit2006,dai2015} and in the T Tauri binary system AS 205 \citep{salyk2014}. There is evidence that our own Solar System might have been affected by a close encounter with another star during its early stages \citep{jilkova2015,pfalzner2018}. It has also been observed that disks in dense regions are generally smaller than the ones in sparser regions \citep{clarke2007,dejuanovelar2012,mann2014}, and it is still up for debate if this is a consequence of dynamical encounters or external photoevaporation, or a combination of both. 

Circumstellar disks are not only affected by external mechanisms, but also by their own viscous growth. The combination of outwardly decreasing angular velocity, together with outwardly increasing angular momentum, causes shearing forces inside the disks. These forces induce the transport of angular momentum from the inner regions, locally lower in angular momentum, to the outer parts. As a consequence mass is accreted from the innermost regions of the disk into its host star, while the outermost regions expand \citep{lynden-bell1974}. Observational reports of spreading disks around T Tauri stars have been described by \citet{isella2009} and \citet{guilloteau2011}. 

From a computational perspective, different approaches have been implemented to assemble these processes together and study their effects on the lifetimes of circumstellar disks. External photoevaporation has been modeled with radiation hydrodynamics codes that solve the flow equations through the disk boundaries, together with photodissociation region (PDR) codes to obtain the temperature profiles of the disks \citep[e.g.][]{haworth2016,facchini2016}. This approach has also been coupled with $\alpha$-disk models to account for viscous spreading \citep[e.g.][]{adams2004,anderson2013,gorti2015,rosotti2017}. The UV radiation fields to which the disks would be subject to in a clustered environment have been estimated from observations and used to calculate the PDR around the hydrodynamic disk. \citet{haworth2018a} introduced the concept of pre-computing photevaporation mass losses in terms of the surface density of the disks, an approach that we expand on in section \ref{photoevaporation}. 

Computational analyses of the impact of dynamical truncations have also been carried out in different ways. The effects of close encounters on a single N-body disk of test particles have been investigated in several studies \citep[e.g.][]{breslau2014,jilkova2016,bhandare2016,pfalzner2018}. In these cases, viscosity is usually neglected since the time scale for a dynamical encounter is much shorter than the viscous time scale for disk spreading. \citet{winter2018,winter2018a} used a ring of test particles around a star to obtain linearized expressions of the effect of a stellar perturber, and then used them to model an SPH disk including viscous spreading. Another approach, in which the dynamics of the cluster are evolved separately and the effects of dynamical encounters over the disks are calculated \textit{a posteriori}, has also been used for this problem \citep[e.g.][]{olczak2006,olczak2010,malmberg2011,steinhausen2014,vincke2015,vincke2016,vincke2018}. Directly adding SPH disks to a simulation of a massive star cluster is extremely computationally expensive. The closest effort corresponds to the work by \citet{rosotti2014}, in which individual SPH codes were coupled to half of the $\SI{1}{M_\odot}$ stars in a small ($N=100$) cluster. This allowed them to study the effects of viscous spreading of the disks and dynamical truncations in a self-consistent way, but were clearly limited by the computational resources needed for this problem. Parametrized approaches have also been developed, where the cluster dynamics and effects of truncations \citep{portegieszwart2016} and viscous spreading \citep{concha-ramirez2019} are considered simultaneously. The truncation of the disks is calculated with expressions obtained from the similarity solutions of \citet{lynden-bell1974}.

In \citet{concha-ramirez2019} we investigated the effect of viscous growth and dynamical truncations on protoplanetary disks inside stars clusters. Both mechanisms were implemented with parametrized models. We showed that considering only viscous evolution and dynamical encounters yields circumstellar disks too large to be comparable to the compact disks observed in star forming regions (in, e.g., the Orion Trapezium cluster \citep{vicente2005,mann2009,robberto2004}, the Lupus clouds \citep{ansdell2018,tazzari2017,alcala2014}, Chamaeleon I \citep{pascucci2016,mulders2017,manara2017}, $\sigma$ Orionis \citep{mauco2016,ansdell2017} and the Upper Scorpio region \citep{barenfeld2017,barenfeld2016}). We argued that other processes must be at play to produce the observed disk size, mass, and accretion rate distributions. In the present work we expand our model by improving our description of the viscous growth of the disks, and by adding external photevaporation to the processes that can shrink them. In this way we look to better represent the environment of star forming regions. 

Instead of the parametrized description used in our previous work, we now model the circumstellar disks using the numerical code VADER \citep{krumholz2015}. This code solves the equations of angular momentum and mass transport in a thin, axisymmetric disk, providing us with a numerical tool to implement the viscous expansion that is less computationally expensive than SPH disks. While the dynamical truncations are still parametrized, the mass loss due to external photevaporation is calculated using the FRIED grid \citep{haworth2018}. This grid consists of pre-calculated mass loss rates for disks of different sizes and masses, immersed in several different external FUV fields. We use the Astrophysical Multipurpose Software Environment \citep[AMUSE\footnote{\url{http://amusecode.org}},][]{portegieszwart2019} framework to bring these codes together along with cluster dynamics and stellar evolution. All the code developed for the simulations, data analyses, and figures of this paper is available in a Github repository\footnote{ZENODO LINK}.


\section{Model}
\label{sec:model}

\subsection{Viscous growth of circumstellar disks}\label{viscous}
We implement the circumstellar disks with viscous evolution using the Viscous Accretion Disk Evolution Resource (VADER) by \citet{krumholz2015}. This is an open source viscous evolution code which uses finite-volume discretization to solve the equations of mass transport, angular momentum, and internal energy in a thin, axisymmetric disk. An AMUSE interface for VADER was developed in the context of this work and is available online\note{add footnote w/ link -- check if Martijn is ok with this!}.  

For the initial disk column density we use the standard disk profile introduced by \citet{lynden-bell1974}:

\begin{equation}\label{profile}
\Sigma(r, t=0) = \Sigma_0 \frac{r_c}{r} \exp\left(\frac{-r}{r_c}\right),
\end{equation}

\noindent
with

\begin{equation}
\Sigma_0 = \frac{m_d}{2 \pi {r_c}^2 \left(1 - \exp\left(-r_d/r_c\right)\right)},
\end{equation}

\noindent
where $r_c$ is the characteristic radius of the disk, $r_d$ and $m_d$ are the initial radius and mass of the disk, respectively, and $\Sigma_0$ is a normalization constant. Considering $r_c \approx r_d$ \citep{anderson2013}, the density profile of the disks takes the form:

\begin{equation}
\Sigma(r, t=0) \approx \frac{m_d}{2 \pi r_d \left(1 - \exp^{-1}\right)} \frac{\exp(-r/r_d)}{r}.
\end{equation}

This expression allows the disk to expand freely at the outer boundary while keeping the condition of zero torque at the inner boundary $r_i$ ($\Sigma(r_i)$ = 0).

To establish the radius of the disks we set the column density outside $r_d$ to a negligible value $\Sigma_{\mathrm{edge}} = \SI{e-12}{\gram\per\square\cm}$. 

The temperature profile of the disks is given by

\begin{equation}
T(r) = T_m \left(\frac{r}{\SI{1}{au}}\right)^{-1/2},
\end{equation}

\noindent
where $T_m$ is the midplane temperature at $\SI{1}{au}$. Based on \citet{anderson2013} we adopt $T_m = \SI{300}{\K}$.

Each disk is composed of a grid of 100 logarithmically spaced cells, in a range between $0.05$ and $\SI{5000}{au}$, with a Keplerian rotation profile. The fact that the size of the grid is much bigger than the disk sizes (which are intially around $\SI{100}{au}$, see section \ref{initdisks}) allows for the disks to expand freely without reaching the border of the grid. The mass flow through the outer boundary of the grid is set to zero in order to maintain the density $\Sigma_{\mathrm{edge}}$ needed to define the disk radius. The mass flow through the inner boundary of the grid is considered as accreted mass and added to the mass of the host star.

\subsection{Dynamical truncations}\label{truncations_dynamical}
Between encounters, circumstellar disks evolve according to the numerical solutions of the transport equations used by VADER. A close encounter with another disk induces a discontinuity in this evolution. To modify the disks we calculate parametrized truncation radii. For two stars of the same mass, \citet{rosotti2014} approximated the truncation radius to a third of the encounter distance. To this we add the mass dependence of \citet{bhandare2016}, to have an expression for an encounter between stars of different masses. The truncation radius then takes the form:

\begin{equation}\label{truncationradius}
{r^\prime} = \frac{r_{enc}}{3}\left(\frac{M}{m}\right)^{0.5},
\end{equation}

\noindent 
where $M$ and $m$ are the masses of the encountering stars.

To implement truncations we first calculate the corresponding truncation radius of the encounter, according to equation \ref{truncationradius}. To define $r^\prime$ as the new disk radius we change the column density of all the disk cells outside $r^\prime$ to the lower value $\Sigma_{\mathrm{edge}}$, as described in section \ref{viscous}.

%Close dynamical encounters not only affect the radius of the affected disks, but also their mass. We followed \citet{portegieszwart2016} to define the amount of mass loss from each disk as:

%\begin{equation}
%dm = m_{d}\frac{r^{1/2} - R^{1/2}}{r^{1/2}}
%\end{equation}

%\noindent
%where $r$ and $R$ are the radii of the disks involved. We calculated the amount of mass that each disk accretes by:

%\begin{equation}\label{accmass}
%dm_{acc} = f \frac{m}{M + m} dm
%\end{equation}

%\noindent
%where $f \leq 1$ is a mass transfer efficiency factor. For our calculations we used $f = 1$. Both equations of mass accretion were applied symmetrically to each of the disks involved in the close encounter.

\subsection{External photoevaporation}\label{photoevaporation}
The amount of mass lost from disks as a result of external photoevaporation depends on the FUV luminosity of the bright stars in the cluster. This luminosity, together with the distance from each of the bright stars to the disks, is used to obtain the amount of radiation received by each disk. Below we explain what we consider to be bright stars and how we calculate the mass loss ratio.


\subsubsection{FUV luminosities}\label{FUVluminosities}
External photoevaporation of circumstellar disks in proximity to OB-type stars is dominated by FUV radiation \citep[see][]{storzer1999,adams2004,pascucci2014,gorti2016,guarcello2016}. We follow \citet{adams2004} in defining the FUV band ranging from 6 eV up to 13.6 eV, or approximately from 912 Å to 2070 Å. We look to calculate the FUV radiation of the stars in the simulated cluster, based on their spectral types. Given the presence of spectral lines in this band, we use synthetic stellar spectra rather than relying on black body approximations. The spectral library used is UVBLUE \citep{rodriguez-merino2005}, chosen for its high coverage of parameter space and high resolution, spanning the appropriate wavelength ranges. To make computations less expensive we use versions of the spectra with a resolving power five times lower than the full ones, which was sufficient for the calculations. 

The UVBLUE spectral library spans a three dimensional parameter space of stellar temperature, metallicity, and surface gravity. We consider all the stars in the cluster to have solar metallicity ($Z = 0.02$). For the temperature and surface gravity, we chose the spectra that most closely resembled the zero age main sequence (ZAMS) values of each star, according to the parametrized stellar evolution code SeBa \citep{portegieszwart1996, toonen2012}.

Using the ZAMS spectra, we calculated the flux and corresponding luminosity using the radius of the star. Given that we know the masses of the stars, we build a relationship between stellar mass and FUV luminosity. This relation takes the shape of a segmented power law, as can be seen in Figure X. A similar fit was obtained by \citet{parravano2003}.
%Ask for more details of this process?

% FUV luminosity vs mass figure!

% Add table?

The mass range of our fit is $0.12 - 100 M_{\odot}$, however, we are only interested in the range $1.9 - 100 M_{\odot}$. As is further explained in section \ref{grid}, we only consider stars with masses higher that $1.9 M_{\odot}$ to be emitting FUV radiation, and $100 M_{\odot}$ is the upper limit for our stellar mass distribution. These bright stars are also subject to stellar evolution, which is implemented with SeBa through the AMUSE interface. The FUV luminosity of each bright star is calculated in every time step, after evolving the stellar evolution code. The small, disked stars were not subject to stellar evolution.

\subsubsection{Mass loss ratio due to external photoevaporation}\label{grid}
To calculate the ratio of mass loss due to external photoevaporation on each disk we use the Far-ultraviolet Radiation Induced Evaporation of Discs (FRIED) grid developed by \citet{haworth2018}. The FRIED grid is an open access grid of calculations of externally evaporating circumstellar disks. It spans a six dimension parameter space consisting of disk sizes ($1 - 400 au$), disk masses ($~10^{-4} - 10^{2} M_{Jup}$), UV fields ($10 - 10^{4} G_0$), stellar masses ($0.05 - 1.9 M_{\odot}$) and disk outer surface densities. The seemingly three dimensional subspace of disk mass, edge density, and disk radius is in fact two dimensional, as any combination of disk radius and disk mass has only one edge density associated with it. Because of this, we only take into account a four dimensional grid of radiation field strength, host star mass, disk mass, and disk radius.

Following the stellar mass ranges of the FRIED grid, we separate the stars in the simulated clusters into two subgroups: \textit{bright stars} and \textit{small stars}. Bright stars are all stars with initial masses higher than $1.9 M_{\odot}$, while small stars have masses lower or equal than $1.9 M_{\odot}$. Only the small stars have disks around them in the simulations. The bright stars are considered as only generating FUV radiation and affecting the disks around the smaller stars. In this way we make sure that we stay within the stellar mass limits of the FRIED grid. Low mass stars ($\lesssim 1 M_\odot$) have a negligible UV flux \citep{adams2006}, so this approximation holds well for our purposes. Calculation of the FUV radiation emitted by the bright stars is further explained in section \ref{FUVluminosities}. In terms of gravitational evolution, both star groups are indistinguishable and they evolve together normally in the gravity code. These star subgroups are considered only for the calculation of FUV radiation and mass loss.

The FRIED grid allows to take a circumstellar disk with a specific mass, size, and density, around a star with a certain mass, and from the FUV radiation that it receives, obtain the photoevaporation mass loss. However, the parameters of the simulated disks not always exactly match the ones in the grid. We perform an interpolation over the grid to calculate the mass losses of the disks in the simulations. Because of computational constraints, we perform the interpolations on a subspace of the grid, such that it contains at least one data point above and below the phase space point of the disk in each dimension. In \note{Appendix X} we show that such a subgrid is enough to perform the interpolation.

\subsubsection{Disk truncation due to photoevaporation}\label{truncations_photoevaporation}
Once the mass loss due to photoevaporation is calculated for every disk, the disks are truncated at a point that coincides with the amount of mass lost in the process. For this, we move outside-in through the disk and remove mass from each cell, by turning its column density to the limit value $\Sigma_{\mathrm{edge}}$. We stop at the point where the mass removed from the disk is equal or slightly less than the calculated mass loss due to photoevaporation. 

We consider the disks to be completely evaporated when they have lost 99\% of their initial mass \citep{anderson2013}. From this point forward the star continues its dynamical evolution normally, but it is not longer affected by FUV radiation from bright stars. 

\subsubsection{Summary of the photoevaporation truncation process}\label{photoevap_summary}
In summary, in every simulation time step, the following procedures are carried out to simulate disk truncation due to photoevaporation. In every time step, for each bright star:

\begin{enumerate}
\item The bright star's FUV luminosity $L_{FUV}$ is calculated as described in section \ref{FUVluminosities}.
\item The distance $d$ from the bright star to each small star is calculated.
\item Using $L_{FUV}$ and $d$, the amount of FUV radiation $\ell_{FUV}$ received by each small star is calculated.
\item Using $\ell_{FUV}$ and the stellar mass, disk mass, and disk size of the small star, we build a subgrid of the FRIED grid and interpolate on it to calculate $\dot{M}_FUV$, the photoevaporation mass loss.
\item The mass evaporated in the time step is subtracted from the disk around the small star.
\end{enumerate}

\subsection{Supernovae explosions}\label{sn}
Because our simulations run for $\SI{10}{Myr}$ (see section \ref{ics}), it is possible that some of the bright stars evolve to the end of their main sequence tracks and explode in a supernova. Hence we introduce a basic handling of these events in our model. Based on \citet{portegieszwart2018} we consider all the stars with disks within $\SI{0.3}{pc}$ to be affected by the explosion. We calculate a truncation radius given by:

\begin{equation}\label{eq_sn}
    r' = 66.02 d^{0.63}  |\cos(i)|^{-0.68}
\end{equation}

\noindent
where $d$ is the distance to the supernova and $i$ is the inclination induced in the disk by the explosion. Since we do not take into account the inclination of the disks, we consider $i = 0$ for all cases.

\subsection{Summary of cluster evolution}
The evolution of the simulated cluster can then be summarized in the following steps. In every time step:

\begin{enumerate}
    \item The dynamics of the cluster are evolved.
    \item Stellar encounters and subsequent disk truncations are handled as explained in section \ref{truncations_dynamical}.
    \item Stellar evolution code is advanced. In case a supernova explosion occurs, the event is handled as defined in section \ref{sn}.
    \item FUV luminosities are calculated for each bright star, and the photoevaporation process is carried out as described in section \ref{photoevap_summary}.
    \item Disk are checked for dispersal. If a disk has been dispersed (see section \ref{truncations_photoevaporation}), the code for the disk is removed and the star continues evolving only as part of the stellar dynamics code.
    \item The simulations run until they reach $\SI{10}{Myr}$ or until all the disks have been dispersed, whichever happens first.
\end{enumerate}

\section{Results}
\label{sec:results}

\subsection{Initial conditions}\label{ics}
\subsubsection{Cluster}
We run X simulations with N stars each located in Plummer (King?) spheres of Z size. We run M simulations lasting for 10 million years each. 

\subsubsection{Disks}\label{initdisks}
The initial sizes of the circumstellar disks are given by:

\begin{equation}
r_d(t=0) = R'\left(\frac{M_*}{M_\odot}\right)^{0.5}
\end{equation}

\noindent
where $R'$ is a constant. The youngest circumstellar disks observed to date have sizes that range from $\sim\SI{30}{au}$ \citep[e.g.][]{lee2018} to $\sim120-\SI{180}{au}$ \citep[e.g.][]{murillo2013,vanthoff2018}. Based on this we choose the value $R' = \SI{100}{au}$, which for our mass range $0.05-1.9 M_\odot$ for stars with disks yields initial sizes between $\sim\SI{22}{au}$ and $\sim\SI{137}{au}$.

\subsection{Photoevaporation vs dynamical truncations (mass loss)}

\subsection{Disk lifetimes}

\subsection{Comparison with observations}

\section{Discussion}
\label{sec:discussion}

\subsection{Disk survival}\label{discussion_disksurvival}

\subsection{Consequences for planet formation}\label{discussion_planetformation}

\subsection{Caveats}\label{caveats}
The present model characterizes a simplified implementation of the processes inside young star forming regions. Several mechanisms and circumstances were not considered for our calculations. Below we discuss how these affect our results.

While photoevaporation is considered to be primarily damaging for disks when coming from external sources, under certain regimes the photons coming from the host star can also contribute to disk dispersal. Internal photoevaporation can form gaps in the protoplanetary disks \note{expand} \citep{gorti2009, gorti2015}. However, for disk regions $> \SI{10}{au}$, external photoevaporation dominates \citep{fatuzzo2008, hollenbach2000}, so the approach used in this work is valid for this approximation. 

The FRIED grid used to estimate the photoevaporative mass loss considers only a 1D model. New simulations by \citet{haworth2019} show that, when considering a 2D disk model, mass loss rates due to external photoevaporation can increase up to a factor 3.7, for the case of a solar mass star whose inital disk is $1\%$ of this mass. The mass loss obtained in this work should then be considered as lower limits, but are still a reasonable estimate of the effects of external photoevaporation.

Circumstellar disks are composed of both dust and gas. Gas disks tend to be larger than dust disks by a factor of $\sim2$ \citep{ansdell2018}. External photoevaporation affects the disk and dust components differently. \note{expand, check dust papers}

In the present work we ignore the existence of binary stars. However, the presence of this or other multiple stellar systems can have direct consequences on both the dynamical evolution of the cluster and on the effects of photoevaporation over the disks. Disks around binary stars have been observed in the star forming regions $\rho$ Ophiuchus \citep{cox2017} and Taurus-Auriga \citep{harris2012,akeson2014,akeson2019}. The observations suggest that these disks are more compact and less bright than the ones around isolated stars. Disks around binary stars might also have shorter lifetimes, due to the effects of the companion on the viscous timescale of the disk and also because of photoevaporation inside the system \citep{shadmehri2018,rosotti2018a}.

Gas leftover in young star forming regions can have important implications on the effect of external photoevaporation \citep{winter2019}. \note{expand}

During the \SI{10}{Myr} of our simulations, it is likely that supernovae explosions occur for the most massive stars. In the present implementation we added a very simplified handling of the effects of supernovae explosions on nearby circumstellar disks, in which they are truncated to a certain radius. In reality the effect of these explosions can have devastating effects that are not instantaneous but maintained over long periods of time. \citet{close2017} show that nearby ($\SI{0.3}{pc}$) supernova explosions can cause mass loss rates of up to $\SI{1E-5}{M_\odot yr^{-1}}$ which can be sustained for about $\SI{200}{yr}$. Only disks that are faced with the flow face-on manage to survive, but still lose $50\%$ of their mass in the process. \citet{portegieszwart2018} show that a supernova explosion at a distance between $0.15$ and $\SI{0.4}{pc}$ could create a misalignment of $\sim 5\degree.6$ between the star and its disk, which is consistent with the inclination of the plane of the Solar System. Such an event would also truncate a disk at around the edge of the Kuiper belt ($42 - \SI{55}{au}$). It is clear that the effects of supernovae can play an important role in the evolution of circumstellar disks in clustered environments. Similar effects can be caused by other outcomes of stellar evolution, such as winds \citep{pelupessy2012}.

Other mechanisms for disk shrinkage should also be considering when modeling them inside young star forming regions. Face-on accretion of gas can make the disks contract as they move through the intracluster medium \citep{wijnen2016, wijnen2017}. \note{expand}

Disk inclinations

The mechanisms driving disk accretion are still up for discussion. The viscosity needed to draw accretion in disks might be due to the viscous disk model, or to magnetohydrodynamical disk winds \citep{ercolano2017}. While some hints of disk winds have been observed, mass loss rates related to this process are still unconstrained due to instrumental restrictions. It is expected that new generations of instruments such as ngVLA will help determine how critical these winds can be to disk evolution \citep{pascucci2018}. 

\section{Summary and conclusions}
\label{sec:summary}


% Don't change these lines
\bsp	% typesetting comment
\bibliographystyle{mnras}
\bibliography{references}
\label{lastpage}
\end{document}

% End of mnras_template.tex
